---
title: "Course Project 1"
author: "Gerardo Reyes Guzman"
date: "23 de diciembre de 2019"
output: html_document
---

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.
Loading and preprocessing the data
We are going to use "mice" and "VIM" packages to explore the data.
1.Code for reading in the dataset and/or processing the data
```{r}
library(readr)
activity <- read_csv("~/activity.csv")
str(activity)
library(mice)
library(VIM)
```
We have 17,568 observations and 3 variables. 
2.Histogram of the total number of steps taken each day
3.Mean and median number of steps taken each day
4.Time series plot of the average number of steps taken
5.The 5-minute interval that, on average, contains the maximum number of steps

```{r}
hist(activity$steps, breaks = 60)
#the histogram shows an activity of cero steps with a frecuency of 10,000. From one to 200 steps the frecuency goes from less than 900 to aproximatly 10 
summary(activity)
#the mean is 37.38 steps; the maximum are 806 steps and there are 2,304 non available observations (NA´s). For the variables "date" and "interval", there are no missing values.
p<-function(x){sum(is.na(x))/length(x)*100}
apply(activity,2,p)
#this 2,304 NA´s corresponds to a 13.11% of the total observations
pairs(activity)
#the maximum number of steps are around 800 and this ocurres three times along the period: mid october, mid november and along december equally distributed
```
6.Code to describe and show a strategy for imputing missing data
7.Histogram of the total number of steps taken each day after missing values are imputed
8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
impute<-mice(activity[,],m=3, seed = 123)
#We use de function "mice" to replace the missing values with three possible values
print(impute)
#describes the number of imputations and the method used to replace the missing values. In this case the method use was the pmm= predictive mean matching
a<-complete(impute,1)
hist(a$steps, breaks = 60)
#Now we plot the histogram with the missing value replaced
pairs(a)
#Now we can see the panel that shows not much difference from the first one. This person does not move quiet a lot, that is why most of the shadow area fall into from zero to two hundred steps.
plot(a$date,a$steps,type="l")
#the graph shows a cyclical behavior of this persons, with a numer of steps falling on weekends and increasing along the working days.
```
What is mean total number of steps taken per day?
the mean is 37.38 steps; the maximum are 806 steps and there are 2,304 non available observations (NA´s). For the variables "date" and "interval", there are no missing values
What is the average daily activity pattern?
the graph shows a cyclical behavior of this persons, with a numer of steps falling on weekends and increasing along the working days.
